### 2021-11-15 documents/sections/extensions.tex
karijona j- review

\subsection{\carijo \obj{j-}}
\label{sec:carijo}
As discussed by \textcite[105--107]{meira1998proto}, \carijo has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-} to \gl{s_a_} verbs.\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.}
This, in combination with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, results in a single unified \gl{s} category \pcref{tab:carreg}.
%
\input{floats/carreg}
%
Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by taking a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}, shown in \exref{car-30}.
%The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.

@@

As noted in \cref{sec:taranoan}, this extension also erased any traces of a potential \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
However, it did not fully eclipse the old \gl{1}\gl{s_a_} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.

\pex<car1w>\carijo \parencite[][5, 42]{guerrero2016karihona}
\a<car-24>
\begingl
\gla wɨ-tə-e=rehe//
\glb \gl{1}-go-\gl{npst}=\gl{frust}//
\glft \qu{I almost go (but I am not going to go).}//
\endgl
\a<car-25>
\begingl
\gla əji-marə-ne w-a-e//
\glb \gl{2}-with-\gl{pl} \gl{1}-be-\gl{npst}//
\glft \qu{I am with you all.}//
\endgl
\xe
%
In addition, the verb \obj{ka} \qu{to say} has a zero-marked first-person form \exref{car-33}.
%We interpret this ∅ as a reflex of \rc{wɨ-}, a result of irregular phonological erosion.
Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would either expect \obj{wɨ-} or \obj{ji-}.
We analyse this ∅ as a reflex of \rc{wɨ-}, primarily based on the (albeit more regular) loss of \rc{w} in \ikpeng and \hixka, and the presumedly higher propensity of an irregular marker for idiosyncratic phonological erosion.
Further, one may suggest the difference between C-initial \obj{ka} \qu{to say} and \obj{tə} \qu{to go} to result from the latter's earlier V-initial nature \pcref{sec:go}.

\ex<car-33>\carijo \perscomm{David Felipe Guerrero-Beltran}\\
\begingl
\glpreamble dëmëmara kae ëwï iya//
\gla n-tə-mə=mara ∅-ka-e əwɨ i-ja//
\glb \gl{3}-go-\gl{pst}=\gl{dub} \gl{1}-say-\gl{npst}.\gl{cert} \gl{1}\gl{pro} \gl{3}-\gl{obl}//
\glft \qu{“Did s/he leave?”, I say to him.}//
\endgl
\xe

>>>>

\subsection{\carijo \obj{j-}}
\label{sec:carijo}
\carijo, the cousin of the Tiriyoan languages, has extended the \gl{1}\gl{s_p_} marker \obj{j(i)-}\footnote{Since all affected \gl{s_a_} verbs are V-initial, only the \envr{}{V} allomorph \obj{j-} occurs in that context.} to \gl{s_a_} verbs \parencite[105--107]{meira1998proto}.
Combined with the extension of \gl{2}\gl{s_a_} \obj{m-} and \gl{1+2}\gl{s_a_} \obj{kɨt-}/\obj{kɨs-} to \gl{s_p_} verbs, this created a single unified \gl{s} category for regularly inflected verbs \pcref{tab:carreg}.
%
\input{floats/carreg}
%
Although the split-\gl{s} system has been lost entirely, former \gl{s_a_} verbs can be identified by the presence of a detransitivizer, like \obj{ehɨnəhɨ} \qu{to fight}, derived from \obj{hɨnəhɨ} \qu{to kill} \parencite[179]{robayo2000avance}, shown in \exref{car-30}.
%The same marker is also found in \gl{3}>\gl{1} scenarios \exref{car1p.car-1} and possessed nouns \exref{car1p.car-31}.

@@

As noted in \cref{sec:taranoan}, this extension also erased any traces of a potential \PTar \gl{1}\gl{s_a_} marker \rc{t-}.
However, it did not fully eclipse the old \gl{1}\gl{s_a_} marker \rc{w-}, which is attested as being preserved in the verbs \obj{tə} \qu{to go} \exref{car1w.car-24} and \obj{a} \qu{to be} \exref{car1w.car-25}.
In addition, the verb \obj{ka} \qu{to say} has a zero-marked first-person form \exref{car1w.car-33}.

\pex<car1w>\carijo \parencite[][5, 42, personal communication]{guerrero2016karihona}
\a<car-24>
\begingl
\gla wɨ-tə-e=rehe//
\glb \gl{1}-go-\gl{npst}=\gl{frust}//
\glft \qu{I almost go (but I am not going to go).}//
\endgl
\a<car-25>
\begingl
\gla əji-marə-ne w-a-e//
\glb \gl{2}-with-\gl{pl} \gl{1}-be-\gl{npst}//
\glft \qu{I am with you all.}//
\endgl
\a<car-33>
\begingl
\glpreamble dëmëmara kae ëwï iya//
\gla n-tə-mə=mara ∅-ka-e əwɨ i-ja//
\glb \gl{3}-go-\gl{pst}=\gl{dub} \gl{1}-say-\gl{npst}.\gl{cert} \gl{1}\gl{pro} \gl{3}-\gl{obl}//
\glft \qu{“Did s/he leave?”, I say to him.}//
\endgl
\xe
%We interpret this ∅ as a reflex of \rc{wɨ-}, a result of irregular phonological erosion.
Based on other C-initial verbs like \obj{tə} \qu{to go} or \obj{tuda} \qu{to arrive}, one would either expect conservative \obj{wɨ-} or innovative \obj{ji-}.
We analyse this ∅ as a reflex of \rc{wɨ-}, primarily based on the (albeit more regular) loss of \rc{w} in \ikpeng and \hixka, and the presumedly higher propensity of an irregular marker for idiosyncratic phonological erosion.
Further, it is possible that the divergent development of \rc{w-} on C-initial \obj{ka} \qu{to say} and \obj{tə} \qu{to go} is a result of the latter's originally V-initial nature \pcref{sec:go}.
